---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Loading and preprocessing the data

````{r}
data <- read.csv(unz("activity.zip", "activity.csv"))
data$date <- as.Date(data$date, format = "%Y-%m-%d")
str(data)
summary(data)
````
## What is mean total number of steps taken per day?

```{r, message=FALSE}
library(dplyr)
library(lubridate) #library for dealing with data
library(ggplot2)
data.day <- data %>% group_by(date) %>% summarize(steps = sum(steps))
steps.mean = mean(data.day$steps,na.rm =TRUE)
steps.median = median(data.day$steps,na.rm =TRUE)
```

The steps mean is `r steps.mean` and the median is `r steps.median`. Here we can see its histogram:

```{r}
hist(data.day$steps[!is.nan(data.day$steps)], xlab = "steps", main = "Histogram of steps", breaks = 30)
```

## What is the average daily activity pattern?
```{r, message=FALSE}
data.interval <- data %>% group_by(interval) %>% summarize(steps = mean(steps, na.rm = TRUE)) 
ggplot(data=data.interval, aes(x=interval, y=steps, na.rm=TRUE))+
  geom_line()+
  geom_smooth()+
  ylim(c(0,205))

```

## Imputing missing values

```{r}
total.na <- sum(is.na(data))
percent.na <- mean(is.na(data))
```

There are `r total.na` NA's in the dataset, which represent the `r percent.na` %.

With the following for loop these missing values are imputed by the mean of the interval.
```{r}
data.filled <- data
for (i in 1:dim(data.filled)[1])
  {
  if (is.na(data.filled$steps[i])== TRUE)
    {
    data.filled$steps[i] <- round( mean(data[data$interval == data.filled[i, "interval"], "steps" ], na.rm = TRUE) )
    }
  
}

```
```{r, message=FALSE}
data.day <- data.filled %>% group_by(date) %>% summarize(steps = sum(steps))
steps.mean = mean(data.day$steps,na.rm =TRUE)
steps.median = median(data.day$steps,na.rm =TRUE)
```

The steps mean is `r steps.mean` and the median is `r steps.median`. Here we can see its histogram:

```{r}
hist(data.day$steps[!is.nan(data.day$steps)], xlab = "steps", main = "Histogram of steps", breaks = 30)
```


## Are there differences in activity patterns between weekdays and weekends?
```{r, message = FALSE}
data.filled$weekend <- ifelse(weekdays(data.filled$date, abbreviate = TRUE) %in% c("sá.", "do."), "weekend", "no weekend")
data.filled$weekend <- as.factor(data.filled$weekend)

data.interval <- data.filled %>% group_by(interval, weekend) %>% summarize(steps = mean(steps, na.rm = TRUE)) 
ggplot(data=data.interval, aes(x=interval, y=steps, fill = weekend))+
  facet_grid(weekend~.)+
  geom_line()+
  geom_smooth()+
  ylim(c(0,205))

```

